---
title: "Confidence Intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{anchoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(designrbi)
```

This article demonstrates how to use the \code{designrbi} package to construct confidence intervals for an average treatment effect (ATE) estimate using a simulation-based approach. We will use data from an anchoring experiment where participants were randomly assigned to either a control or treatment group, and we will calculate the ATE, simulate experiments under the null hypothesis, and construct a confidence interval.

The complete pipeline to compute a 95% confidence interval for the ATE estimate is as follows:

```{r eval=FALSE}
anchor |>
  experiment_to_schedule(treatment = d, response = y) |>
  impute_unobserved(tau = ate_hat) |>
  sim_experiment(reps = 5000) |>
  group_by(replicate) |>
  summarize(ate_hat = diff_in_means(y, d)) |>
  quantile(probs = c(0.025, 0.975))
```

This article will walk through each step of this pipeline in detail, starting with understanding the experimental data, then imputing a schedule of potential outcomes, simulating new experiments, and finally calculating the confidence interval.

### Understanding the experimental data

Start by loading the necessary libraries and the data.

```{r message = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
anchor <- read_csv("https://stat158.berkeley.edu/spring-2026/data/anchoring/anchoring.csv") |>
  mutate(d = factor(X),
         y = Y) |>
  select(d, y)
anchor
```

Note the structure of experimental data: each row corresponds to a single unit (participant) and contains the treatment assignment (\code{d}) and the observed response (\code{y}).

In this experiment, the estimand of interest is the average treatment effect (ATE), which can be estimated using the difference in means between the treatment and control groups. We can write such a function to operate on the response and treatment vectors and use it to calculate the observed ATE estimate.

```{r}
diff_in_means <- function(y, x) {
  groups <- split(y, x)
  Ybarhat0 <- mean(groups[[1]])
  Ybarhat1 <- mean(groups[[2]])
  Ybarhat1 - Ybarhat0
}

ate_hat <- diff_in_means(anchor$y, anchor$d)
ate_hat
```


### Impute a schedule of potential outcomes

A RBI confidence interval seeks to quantify the variability that was induced by the random assignment that occurred at the outset of the experiment. To do this, we first need to impute the unobserved potential outcomes for each unit in the experiment. Let's start by creating a schedule of the observed potential outcomes.

```{r}
schedule <- anchor |>
  experiment_to_schedule(treatment = d, response = y)
schedule
```

Note that the schedule contains the observed potential outcomes for each unit, but the unobserved potential outcomes are missing. We can use the observed ATE estimate to impute the unobserved potential outcomes by assuming that the treatment effect is constant across all units.

```{r}
imputed_schedule <- impute_unobserved(schedule, tau = ate_hat)
imputed_schedule
```

While this is likely not the true schedule of potential outcomes, it is a useful tool for simulating the random assignment process and understanding the variability of our ATE estimate.


### Simulate a single experiment

Now that we have a schedule of potential outcomes, we can simulate the random assignment process to generate new datasets that could have been observed under the same experimental design. Let's start by simulating a single experiment.

```{r}
set.seed(402) # for reproducibility
sim1 <- imputed_schedule |>
  sim_experiment(reps = 1)
sim1
```

We can compare this to the original data set:

```{r}
anchor
```

The units are the same across both data sets, but the treatment assignments and observed responses may differ due to the random assignment process. The first unit, for example was assigned to `73` simulated experiment, so we observed $Y_1(1)$: 31.5. In the original data, the first unit was assigned to `11` and we observed $Y_1(0)$: 16. The second unit, was assigned to `11` in both the original experiment and the simulated experiment, so we observed $Y_2(0)$: 5 in both cases.

We can calculate the ATE estimate for this simulated experiment.

```{r}
ate_hat_sim1 <- sim1 |>
  summarize(ate_hat = diff_in_means(y, d))
ate_hat_sim1
```

This serves as a proof of concept that we can simulate new experiments and calculate new ATE estimates under the same experimental design. However, a single simulated experiment is not enough to understand the variability of our ATE estimate. We need to simulate many experiments to get a distribution of ATE estimates.


### Simulating many experiments

Let's use the same approach but simulate 5000 experiments.

```{r}
sim5000 <- imputed_schedule |>
  sim_experiment(reps = 5000)
head(sim5000)
tail(sim5000)
```

`sim_experiment()` returns the 5000 simulated experiments as a single tall data frame, each one stacked on top of one another, with a new variable called `replicate` that indicates which rows belong to which simulated experiment. We can use this variable to group the data and calculate the ATE estimate for each simulated experiment.

```{r}
sim5000_estimates <- sim5000 |>
  group_by(replicate) |>
  summarize(ate_hat = diff_in_means(y, d))
sim5000_estimates
```

We can visualize the distribution of ATE estimates from the simulated experiments and see that it is centered around the observed ATE estimate from the original experiment.
```{r}
ggplot(sim5000_estimates, aes(x = ate_hat)) +
  geom_histogram(binwidth = 0.5, color = "black") +
  geom_vline(xintercept = ate_hat, color = "red", linetype = "dashed") +
  labs(title = "Distribution of Estimates from Simulated Experiments",
       x = "Estimated ATE",
       y = "Frequency") +
  theme_bw()
```

### Calculate confidence interval

A $1-\alpha$ confidence interval summarizes the distribution of ATE estimates from the simulated experiments with two numbers, the lower and upper bounds, taken from the quantiles of the distribution.

```{r}
alpha <- 0.05 # for a 95% confidence interval
ci <- quantile(sim5000_estimates$ate_hat, probs = c(alpha/2, 1 - alpha/2))
ci
```

