---
title: "Simulate an Experiment from a Schedule"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{anchoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(designrbi)
```

This article demonstrates how to simulate an experiment from a pre-existing schedule of potential outcomes. This is useful for calculating power or quantifying the coverage probability of confidence intervals under a particular data generating process.

### Create a schedule object

The key function for simulating an experiment is \code{sim_experiment()}. This function takes as input a *schedule* object. These objects can be created using `experiment_to_schedule()` if you have an observed dataset with a treatment variable and an outcome variable. If instead you have a data frame that contains the potential outcomes for each unit under each treatment condition, you can use `as_schedule()` to convert it to a schedule object.

Let's start with an example schedule:

```{r}
library(readr)

anchoring_sched <- read_csv("https://stat158.berkeley.edu/spring-2026/data/anchoring/godlike_schedule_A.csv")
anchoring_sched$d <- factor(anchoring_sched$d)
anchoring_sched
```

This data frame contains the potential outcomes for 62 units in the Anchoring Experiment, where they were assigned to answer a question that referenced the number `11` or `73` and then we asked to estimated a percentage of African countries in the United Nations. The columns `Y11` and `Y73` represent the potential outcomes for each unit under the control and treatment conditions, respectively. Note these are not the observed outcomes, but rather the potential outcomes that would be observed under each treatment condition.

This data frame also contains a factor column that indicates the treatment assignments. This is useful because it represents the total number of units that were assigned to each group, and essential piece of information for simulating experiments that mirror the original design.

We can convert this data frame to a schedule object using `as_schedule()`.

```{r}
schedule <- as_schedule(anchoring_sched, treatment = d, potential_outcomes = c(Y11, Y73))
schedule
```

### Simulate an experiment from the schedule

Now that we have a schedule object, we can simulate an experiment using the `sim_experiment()` function. This function takes as input a schedule object and the number of simulations to run.

```{r}
set.seed(5024) # for reproducibility
simulated_experiment <- schedule |>
  sim_experiment(reps = 1)
simulated_experiment
```

This data frame contains the observed outcomes for each unit under the simulated treatment assignments. Each row corresponds to a single unit, and the columns represent the treatment assignment and the observed outcome for that unit in the simulated experiment.
